---
title: "CRF maps"
author: "Mark Taylor"
date: "21/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# load packages
library(tidyverse)
library(readxl)
library(scales)
library(janitor)
library(ggrepel)
library(ggridges)
library(ggforce)
library(sf)
library(plotly)

# load everything that isn't a shapefile

# crf money: from the ace site
crf_money_round1 <- 
  read_excel("CRF investment data published 020421.xlsx", 
             2,
             .name_repair = "universal") %>% 
  select(-Strand) %>% 
  mutate(source = "ACE grants round 1")
crf_money_capitalkickstart <- 
  read_excel("CRF investment data published 020421.xlsx", 
             3,
             .name_repair = "universal") %>% 
  mutate(source = "ACE capital kickstart")
crf_money_round2 <- 
  read_excel("CRF investment data published 020421.xlsx", 
             4,
             .name_repair = "universal") %>% 
  mutate(source = "ACE grants round 2")

# crf money: from dcms
# https://www.gov.uk/government/news/400-million-to-help-more-than-2700-arts-culture-heritage-organisations-and-independent-cinemas-survive-and-thrive
# sheet 1 seems to be the same, so steering clear for the moment
# bfi: weird missingness, trying to deal by replacing NAs with "per cinema"
bfi_money <- 
  read_excel("CRF_2-Awards_read-only.xlsx", 
             2, 
             skip = 1,
             na = "-",
             .name_repair = "universal") %>% 
  mutate(Award.... = 
           coalesce(Award...., 
                    Award.per.Cinema....))
# nlhf: looks ok
nlhf_money <- 
  read_excel("CRF_2-Awards_read-only.xlsx", 3, 
             skip = 1,
             .name_repair = "universal")

# repayable finance from dcms - seems fine
repayable_finance <- 
  read_excel("CRF_2-Awards_read-only.xlsx", 4, 
             skip = 1,
             .name_repair = "universal") %>% 
  mutate(Local.authority = fct_recode(Local.authority,
                                        "York" = 
                                          "City of York"))


# population data from 2019
# has updated local authority codes
# https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatesforukenglandandwalesscotlandandnorthernireland
population <- 
  read_excel("ukmidyearestimates20192020ladcodes.xls", 
             6,
             skip = 4,
             .name_repair = "universal") %>% 
  select(Code:All.ages) %>% 
  rename(la_name = 
           Name,
         population = 
           All.ages) %>% 
  filter(Geography1 != "Country" &
           Geography1 != "Region" &
           Geography1 != "Metropolitan County" &
           Geography1 != "County"
  ) %>% 
  mutate(country = substr(Code, 
                          1,
                          1)) %>% 
  filter(country == "E") %>% 
  select(-country)

# combine crf money
crf_money <- 
  bind_rows(crf_money_round1,
            crf_money_round2,
            crf_money_capitalkickstart)

# load shapefile
local_authorities <- 
  read_sf(dsn = "shapefiles/Local_Authority_Districts_(December_2020)_UK_BUC", 
  layer = "Local_Authority_Districts_(December_2020)_UK_BUC")

# ok, need a data frame with a column for all sources of money
all_money_sources <- 
  crf_money %>% 
  group_by(Local.Authority,
           source) %>% 
  summarise(money = 
              sum(..Awarded)) %>% 
  pivot_wider(names_from = "source",
              values_from = "money") %>% 
  full_join(bfi_money %>% 
    rename(Local.Authority = 
             Local.authority) %>% 
    group_by(Local.Authority) %>% 
    summarise(BFI  = 
                sum(Award....))) %>% 
    full_join(nlhf_money %>% 
      rename(Local.Authority = 
               Local.authority) %>% 
      group_by(Local.Authority) %>% 
      summarise(NLHF  = 
                  sum(Award....))) %>% 
  pivot_longer(-Local.Authority) %>% 
  mutate(value = replace_na(value, 0)) %>% 
  rename(Source = name,
         Money = value)

# combine with population data
all_money_population <- 
  all_money_sources %>% 
  full_join((population %>% 
               rename(Local.Authority = 
                        la_name))) %>% 
  mutate(money_per_head = 
           round((Money / 
                population), 
                2))

# combine with shapefile, limit to england
money_population_geography <- 
  full_join(all_money_population, 
            (local_authorities %>% 
               rename(Code = 
                        LAD20CD)) )%>% 
  mutate(country = substr(Code, 
                          1,
                          1)) %>% 
  filter(country == "E") %>% 
  select(-country)
```

## What is this document?

This is a home for a load of maps I've made about where the CRF money has gone. It's basically the same information as in the other document, just in map format.

If it doesn't work, try it again in Chrome; if that doesn't work, chase me. You should be able to zoom in on the maps, as well as hover over them, and find out which Local Authority you're pointing at, how much money they got overall from each source, and how much per head. Maps are coloured in according to the total amount of money that went to each LA, not per head (as doing it that way everything gets warped by the City of London).

## ACE grants: round one

```{r}

ace_round1 <- 
  money_population_geography %>%
  filter(Source == "ACE grants round 1") 

for_ggplotly_ace_round1 <- 
  ggplot(ace_round1) +
  aes(geometry = geometry,
      fill = Money,
      text = paste(Local.Authority, 
                   ": \n£", 
                   money_per_head, 
                   "per head \n Total: £", 
                   comma(Money, 
                         accuracy = 2L))) +
  geom_sf() +
  theme_void() +
  scale_fill_viridis_c(direction = -1,
                       option = "A") +
  theme(legend.position = "none") 

ggplotly(for_ggplotly_ace_round1,
         tooltip = "text")
```

## ACE grants: round two

```{r}

ace_round2 <- 
  money_population_geography %>%
  filter(Source == "ACE grants round 2") 

for_ggplotly_ace_round2 <- 
  ggplot(ace_round2) +
  aes(geometry = geometry,
      fill = Money,
      text = paste(Local.Authority, 
                   ": \n£", 
                   money_per_head, 
                   "per head \n Total: £", 
                   comma(Money, 
                         accuracy = 2L))) +
  geom_sf() +
  theme_void() +
  scale_fill_viridis_c(direction = -1,
                       option = "A") +
  theme(legend.position = "none") 

ggplotly(for_ggplotly_ace_round2,
         tooltip = "text")
```

## ACE capital kickstart

```{r}


ace_capital_kickstart_plot_data <- 
  money_population_geography %>%
  filter(Source == "ACE capital kickstart") 

for_ggplotly_ace_capital_kickstart <- 
  ggplot(ace_capital_kickstart_plot_data) +
  aes(geometry = geometry,
      fill = Money,
      text = paste(Local.Authority, 
                   ": \n£", 
                   money_per_head, 
                   "per head \n Total: £", 
                   comma(Money, 
                         accuracy = 2L))) +
  geom_sf() +
  theme_void() +
  scale_fill_viridis_c(direction = -1,
                       option = "A") +
  theme(legend.position = "none") 

ggplotly(for_ggplotly_ace_capital_kickstart,
         tooltip = "text")
```

## BFI

```{r}

bfi_plot_data <- 
  money_population_geography %>%
  filter(Source == "BFI") 

for_ggplotly_bfi <- 
  ggplot(bfi_plot_data) +
  aes(geometry = geometry,
      fill = Money,
      text = paste(Local.Authority, 
                   ": \n£", 
                   money_per_head, 
                   "per head \n Total: £", 
                   comma(Money, 
                         accuracy = 2L))) +
  geom_sf() +
  theme_void() +
  scale_fill_viridis_c(direction = -1,
                       option = "A") +
  theme(legend.position = "none") 

ggplotly(for_ggplotly_bfi,
         tooltip = "text")
```

## NLHF

```{r}

nlhf <- 
  money_population_geography %>%
  filter(Source == "NLHF") 

for_ggplotly_nlhf <- 
  ggplot(nlhf) +
  aes(geometry = geometry,
      fill = Money,
      text = paste(Local.Authority, 
                   ": \n£", 
                   money_per_head, 
                   "per head \n Total: £", 
                   comma(Money, 
                         accuracy = 2L))) +
  geom_sf() +
  theme_void() +
  scale_fill_viridis_c(direction = -1,
                       option = "A") +
  theme(legend.position = "none") 

ggplotly(for_ggplotly_nlhf,
         tooltip = "text")
```