---
title: "Money by engagement, and constituency type"
author: "Mark Taylor"
date: "11/05/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# load packages
library(tidyverse)
library(readxl)
library(scales)
library(janitor)
library(ggrepel)
library(ggridges)
library(ggforce)
library(sf)
library(plotly)
library(DT)
library(kableExtra)


# crf money: from the ace site
crf_money_round1 <- 
  read_excel("CRF investment data published 020421.xlsx", 
             2,
             .name_repair = "universal") %>% 
  select(-Strand) %>% 
  mutate(source = "ACE grants round 1")
crf_money_capitalkickstart <- 
  read_excel("CRF investment data published 020421.xlsx", 
             3,
             .name_repair = "universal") %>% 
  mutate(source = "ACE capital kickstart")
crf_money_round2 <- 
  read_excel("CRF investment data published 020421.xlsx", 
             4,
             .name_repair = "universal") %>% 
  mutate(source = "ACE grants round 2")

# crf money: from dcms
# https://www.gov.uk/government/news/400-million-to-help-more-than-2700-arts-culture-heritage-organisations-and-independent-cinemas-survive-and-thrive
# sheet 1 seems to be the same, so steering clear for the moment
# bfi: weird missingness, trying to deal by replacing NAs with "per cinema"
bfi_money <- 
  read_excel("CRF_2-Awards_read-only.xlsx", 
             2, 
             skip = 1,
             na = "-",
             .name_repair = "universal") %>% 
  mutate(Award.... = 
           coalesce(Award...., 
                    Award.per.Cinema....))
# nlhf: looks ok
nlhf_money <- 
  read_excel("CRF_2-Awards_read-only.xlsx", 3, 
             skip = 1,
             .name_repair = "universal")

# repayable finance from dcms - seems fine
repayable_finance <- 
  read_excel("CRF_2-Awards_read-only.xlsx", 4, 
             skip = 1,
             .name_repair = "universal") %>% 
  mutate(Local.authority = fct_recode(Local.authority,
                                        "York" = 
                                          "City of York"))


# population data from 2019
# has updated local authority codes
# https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatesforukenglandandwalesscotlandandnorthernireland
population <- 
  read_excel("ukmidyearestimates20192020ladcodes.xls", 
             6,
             skip = 4,
             .name_repair = "universal") %>% 
  select(Code:All.ages) %>% 
  rename(la_name = 
           Name,
         population = 
           All.ages) %>% 
  filter(Geography1 != "Country" &
           Geography1 != "Region" &
           Geography1 != "Metropolitan County" &
           Geography1 != "County"
  ) %>% 
  mutate(country = substr(Code, 
                          1,
                          1)) %>% 
  filter(country == "E") %>% 
  select(-country)

# combine crf money
crf_money <- 
  bind_rows(crf_money_round1,
            crf_money_round2,
            crf_money_capitalkickstart)


npo_data <- 
  read_excel("NPO_2018_12082020_0.xlsx") %>% 
  mutate(Organisation = 
           `Applicant Name`)

npo_names <- 
  npo_data %>% 
  select(Organisation) %>% 
  pull()

# add flag to crf money
# of whether it's an npo or not

crf_money_capitalkickstart <- 
  crf_money_capitalkickstart %>% 
  mutate(is_npo = 
           if_else(Organisation %in% npo_names,
                   "NPO",
                   "Not NPO")) 

crf_money_round1 <- 
  crf_money_round1 %>% 
  mutate(is_npo = 
           if_else(Organisation %in% npo_names,
                   "NPO",
                   "Not NPO")) 
crf_money_round2 <- 
  crf_money_round2 %>% 
  mutate(is_npo = 
           if_else(Organisation %in% npo_names,
                   "NPO",
                   "Not NPO")) 


all_money_sources <- 
  crf_money %>% 
  group_by(Local.Authority,
           source) %>% 
  summarise(money = 
              sum(..Awarded)) %>% 
  pivot_wider(names_from = "source",
              values_from = "money") %>% 
  full_join(bfi_money %>% 
    rename(Local.Authority = 
             Local.authority) %>% 
    group_by(Local.Authority) %>% 
    summarise(BFI  = 
                sum(Award....))) %>% 
    full_join(nlhf_money %>% 
      rename(Local.Authority = 
               Local.authority) %>% 
      group_by(Local.Authority) %>% 
      summarise(NLHF  = 
                  sum(Award....))) %>% 
  pivot_longer(-Local.Authority) %>% 
  mutate(value = replace_na(value, 0)) %>% 
  rename(Source = name,
         Money = value)

# combine with population data
all_money_population <- 
  all_money_sources %>% 
  full_join((population %>% 
               rename(Local.Authority = 
                        la_name))) %>% 
  mutate(money_per_head = 
           round((Money / 
                population), 
                2))

# add active lives data
active_lives <- 
  read_csv("active_lives.csv")

# add constituency type data
hoc <- 
  read_csv("pcon-classification-csv.csv")

# single row for each constituency
# just England
hoc_use <- 
  hoc %>% 
  group_by(constituency_name) %>% 
  mutate(population = sum(population)) %>% 
  top_n(1, percent_of_constituency) %>% 
  filter(str_sub(constituency_code,
                 1, 1) == "E")


crf_constituency <- 
  crf_money %>% 
  group_by(Constituency) %>% 
  summarise(total_money = sum(..Awarded)) %>% 
  rename(constituency_name = 
           Constituency)

```

## CRF funding and a composite measure of arts engagement

In this plot, I've combined data from the Cultural Recovery Fund (specifically, combining the different pots distributed by ACE, but *not* the pots distributed by NLHF or BFI), with data from Active Lives, which is a survey that includes at least 500 people per Local Authority, allowing estimates of cultural activity per LA.

Please note that the x-axis, which reports total CRF funding per head, is on a log scale: that is, the space between £1 and £10 per head is the same as £10 and £100 per head. This is so that we don't end up with the vast majority of LAs hemmed into the left-hand part of the plot. 

Please also note that the measure of cultural activity we're using here is the percentage of people per LA who said that they'd done at least one of "Spent time doing a creative, artistic, theatrical or music activity or craft", or "Attended an event, performance or festival involving creative, artistic, dance, theatrical or music activity". It's weird that I can't get just one or the other -- dunno why this is, sorry.

```{r cars}
money_active_lives <- 
  (all_money_population %>% 
  group_by(Local.Authority) %>% 
  summarise(total_money_per_head = 
              sum(money_per_head)) %>% 
  full_join(active_lives %>% 
               rename(Local.Authority = 
                        LA_name) %>% 
              mutate(Local.Authority = 
                       fct_recode(Local.Authority,
                                  "King's Lynn and West Norfolk" = 
                                    "Kings Lynn and West Norfolk"))) %>% 
  filter(Local.Authority != "City of London") %>% 
  ggplot() +
  aes(x = percent_culture,
      y = total_money_per_head,
      text = paste(Local.Authority, 
                   ": \n£", 
                   total_money_per_head, 
                   "per head \n ",
                   round(percent_culture*100),
                   "% adults attending or participating in culture \n (Active Lives, 2018-19)")) +
  geom_point() +
  scale_y_log10() +
    theme_minimal() +
    scale_x_continuous(labels = percent) +
    labs(x = "% attending or participating in culture (Active Lives 2018-19)",
         y = "Total CRF £ per head (all sources)"))

ggplotly(money_active_lives,
         tooltip = "text")
```

This shows us that there's a relationship between LA-level cultural activity, and CRF funding, but it's not particularly strong.

## CRF and different classifications of place

I've drawn a bunch of maps about where CRF money has gone, but they're often not particularly easy to interpret; more densely-populated areas are less visible on maps, it can be difficult to follow patterns, and so on. It's also often not that helpful to include information about things like the top 10 local authorities, etc, as the difference between eleventh place and last place -- even if not as large in £ terms as the difference between first and tenth place -- is likely to be significant across the entire group.

Because of this, I'm using the [House of Commons Library classification of constituencies and local authorities](https://commonslibrary.parliament.uk/research-briefings/cbp-8322/). You can read more about this at the link, but the idea's to group together areas that are similar to each other. Terms like "town" and "city" aren't used in a formal, cathedral-focused way -- Reading's classified as a city, York as a town -- but in relation to places' populations. The categories are, with numbers in brackets for how many there are across the UK:

- Core City (London)
- Core City (outside London: 11): there are eleven of these, such as Glasgow, Sheffield, and Bristol
- Other City (24): places with more than 175,000 people, such as Leicester, Portsmouth, and Aberdeen
- Large Town (119): places with 60,000-175,000 people, such as Rotherham, Cambridge, and Maidstone
- Medium Town (270): places with 25,000-60,000 people, such as Gravesend, Jarrow, and Exmouth
- Small Town (674): places with 7,5000-25,000 people, such as Falmouth, New Romney, and Holbeach
-  Village (6,116): places with less than 7,500 people

Here, I've used the parliamentary constituency data from the CRF tables. This means that constituencies are classified based on whichever category has the largest fraction of the population in it. For example, 61% of people in the Macclesfield parliamentary constituency are classified as living in a medium town (Macclesfield itself), while 19% are classified as living in a small town (like Poynton) and 20% in a village (like Bollington). In this case, it means if an organisation in Poynton has had CRF money, it'll be classified as being part of a medium town, even though ideally we'd classify it as being in a small town.

Overall, the classification by constituency is (this is just England):
- Core City (London): 73
- Core City (outside London): 41
- Other city: 54
- Large town: 124
- Medium town: 100
- Small town: 48
- Village or smaller: 93

First, let's look at the overall money per head to each of these different categories:

```{r}

hoc_use %>% 
  full_join(crf_constituency) %>% 
  mutate(total_money = replace_na(total_money, 0)) %>% 
  group_by(classification) %>% 
  summarise(total_population = sum(population),
            total_money = sum(total_money)) %>% 
  mutate(money_per_head = 
           total_money / 
           total_population) %>% 
  na.omit %>% 
  mutate(classification = fct_relevel(classification,
                                      "Core City (London)",
                                      "Core City (outside London)",
                                      "Other City",
                                      "Large Town",
                                      "Medium Town",
                                      "Small Town",
                                      "Village or smaller")) %>% 
  ggplot() +
  aes(x = money_per_head,
      y = fct_rev(classification),
      label = paste0("£", round(money_per_head, 2))) +
  geom_col() +
  geom_text(hjust = 0,
            nudge_x = .5) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_blank()) +
  labs(x = "",
       y = "")
```

This shows us that the largest spend per head has gone to the non-London core cities, followed by London. To an extent, this will reflect the different ways that constituencies are classified. Constituencies in the outer boroughs of London are classified as being in London, while constituencies in the city-regions of the core cities, but outside the LA boundaries of those cities themselves, aren't; for example, Salford  & Eccles is classified as a large town under this classifications. My guess is that if we separated out inner and outer London, we'd see much higher spend per head in inner than in outer.

On the other end, the areas with the least money spent on them are medium towns -- even less than small towns, and villages and smaller. We might have expected that people in the areas with the largest populations would have the most money spent on them, and that this would gradually decrease as populations decrease, but this isn't the case.

Next, let's look at how this funding has varied across the different pots, and how it's split between NPOs and non-NPOs:

```{r}

hoc_use %>% 
  full_join((crf_money_round1 %>% 
               rename(constituency_name = 
                        Constituency))) %>% 
  mutate(..Awarded = 
           replace_na(..Awarded, 0)) %>% 
  group_by(classification,
           source,
           is_npo) %>% 
  summarise(total_money = sum(..Awarded)) %>% 
  na.omit %>% 
  bind_rows(hoc_use %>% 
              full_join((crf_money_round2 %>% 
                           rename(constituency_name = 
                                    Constituency))) %>% 
              mutate(..Awarded = 
                       replace_na(..Awarded, 0)) %>% 
              group_by(classification,
                       source,
                       is_npo) %>% 
              summarise(total_money = sum(..Awarded)) %>% 
              na.omit) %>% 
  bind_rows(hoc_use %>% 
              full_join((crf_money_capitalkickstart %>% 
                           rename(constituency_name = 
                                    Constituency))) %>% 
              mutate(..Awarded = 
                       replace_na(..Awarded, 0)) %>% 
              group_by(classification,
                       source,
                       is_npo) %>% 
              summarise(total_money = sum(..Awarded)) %>% 
              na.omit) %>% 
  full_join(hoc_use %>% 
              group_by(classification) %>% 
              summarise(population = sum(population))) %>% 
  mutate(money_per_head = 
           total_money / 
           population) %>% 
  mutate(source = fct_relevel(source,
                              "ACE grants round 1",
                              "ACE grants round 2",
                              "ACE capital kickstart")) %>% 
  ggplot() +
  aes(x = money_per_head,
      y = fct_rev(fct_relevel(classification,
                              "Core City (London)",
                              "Core City (outside London)",
                              "Other City",
                              "Large Town",
                              "Medium Town",
                              "Small Town",
                              "Village or smaller")),
      label = paste0("£", round(money_per_head, 2))) +
  geom_col() +
  geom_text(hjust = 0,
            nudge_x = .5) +
  facet_grid(is_npo ~ source) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_blank()) +
  theme(panel.spacing = unit(1, "lines")) +
  labs(x = "",
       y = "") + 
  coord_cartesian(clip = "off")
```

Hidden in this graph is The Factory taking up £4:71 of the £5:81 per head spent in ACE capital kickstart money to non-NPOs in Core Cities.

Other than that, we can see that there's significant differences in how much money went to NPOs across the different pots across different areas; while the differences are also pronounced for non-NPOs, they're not as proportionally large. Also interesting is the fact that London got the most money per head among non-NPOs, but other core cities got significantly more money per head among NPOs.