---
title: "Where did the money go?"
author: "Mark Taylor"
date: "22/04/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE)
# load packages
library(tidyverse)
library(readxl)
library(scales)
library(janitor)
library(ggrepel)
library(ggridges)
library(ggforce)
library(sf)
library(plotly)
library(DT)
library(kableExtra)

# load everything that isn't a shapefile

# crf money: from the ace site
crf_money_round1 <- 
  read_excel("CRF investment data published 020421.xlsx", 
             2,
             .name_repair = "universal") %>% 
  select(-Strand) %>% 
  mutate(source = "ACE grants round 1")
crf_money_capitalkickstart <- 
  read_excel("CRF investment data published 020421.xlsx", 
             3,
             .name_repair = "universal") %>% 
  mutate(source = "ACE capital kickstart")
crf_money_round2 <- 
  read_excel("CRF investment data published 020421.xlsx", 
             4,
             .name_repair = "universal") %>% 
  mutate(source = "ACE grants round 2")

# crf money: from dcms
# https://www.gov.uk/government/news/400-million-to-help-more-than-2700-arts-culture-heritage-organisations-and-independent-cinemas-survive-and-thrive
# sheet 1 seems to be the same, so steering clear for the moment
# bfi: weird missingness, trying to deal by replacing NAs with "per cinema"
bfi_money <- 
  read_excel("CRF_2-Awards_read-only.xlsx", 
             2, 
             skip = 1,
             na = "-",
             .name_repair = "universal") %>% 
  mutate(Award.... = 
           coalesce(Award...., 
                    Award.per.Cinema....))
# nlhf: looks ok
nlhf_money <- 
  read_excel("CRF_2-Awards_read-only.xlsx", 3, 
             skip = 1,
             .name_repair = "universal")

# repayable finance from dcms - seems fine
repayable_finance <- 
  read_excel("CRF_2-Awards_read-only.xlsx", 4, 
             skip = 1,
             .name_repair = "universal") %>% 
  mutate(Local.authority = fct_recode(Local.authority,
                                        "York" = 
                                          "City of York"))


# population data from 2019
# has updated local authority codes
# https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatesforukenglandandwalesscotlandandnorthernireland
population <- 
  read_excel("ukmidyearestimates20192020ladcodes.xls", 
             6,
             skip = 4,
             .name_repair = "universal") %>% 
  select(Code:All.ages) %>% 
  rename(la_name = 
           Name,
         population = 
           All.ages) %>% 
  filter(Geography1 != "Country" &
           Geography1 != "Region" &
           Geography1 != "Metropolitan County" &
           Geography1 != "County"
  ) %>% 
  mutate(country = substr(Code, 
                          1,
                          1)) %>% 
  filter(country == "E") %>% 
  select(-country)

# combine crf money
crf_money <- 
  bind_rows(crf_money_round1,
            crf_money_round2,
            crf_money_capitalkickstart)

# load shapefile
local_authorities <- 
  read_sf(dsn = "shapefiles/Local_Authority_Districts_(December_2020)_UK_BUC", 
  layer = "Local_Authority_Districts_(December_2020)_UK_BUC")

# ok, need a data frame with a column for all sources of money
all_money_sources <- 
  crf_money %>% 
  group_by(Local.Authority,
           source) %>% 
  summarise(money = 
              sum(..Awarded)) %>% 
  pivot_wider(names_from = "source",
              values_from = "money") %>% 
  full_join(bfi_money %>% 
    rename(Local.Authority = 
             Local.authority) %>% 
    group_by(Local.Authority) %>% 
    summarise(BFI  = 
                sum(Award....))) %>% 
    full_join(nlhf_money %>% 
      rename(Local.Authority = 
               Local.authority) %>% 
      group_by(Local.Authority) %>% 
      summarise(NLHF  = 
                  sum(Award....))) %>% 
  pivot_longer(-Local.Authority) %>% 
  mutate(value = replace_na(value, 0)) %>% 
  rename(Source = name,
         Money = value)

# combine with population data
all_money_population <- 
  all_money_sources %>% 
  full_join((population %>% 
               rename(Local.Authority = 
                        la_name))) %>% 
  mutate(money_per_head = 
           round((Money / 
                population), 
                2))

# combine with shapefile, limit to england
money_population_geography <- 
  full_join(all_money_population, 
            (local_authorities %>% 
               rename(Code = 
                        LAD20CD)) )%>% 
  mutate(country = substr(Code, 
                          1,
                          1)) %>% 
  filter(country == "E") %>% 
  select(-country)


# load npo data
npo_data <- 
  read_excel("NPO_2018_12082020_0.xlsx") %>% 
  mutate(Organisation = 
           `Applicant Name`)

# npo or not

npo_names <- 
  npo_data %>% 
  select(Organisation) %>% 
  pull()

crf_money <- 
  crf_money %>% 
  mutate(is_npo = 
           if_else(Organisation %in% npo_names,
                   "NPO",
                   "Not NPO")) 
```

## Identifying specific recipients within local authorities

How come local authorities like Manchester are home to such a large amount of Cultural Recovery Fund money? Is it because there's loads of institutions there that received medium-sized pots of money, or because there's a small number who received massive pots of money (that don't just work within that local authority, like the Science Museum Group)? Let's build a widget so people can explore. You can also see whether each organisation getting money is an NPO or not.

```{r}
crf_money %>% 
  rename(Grant = 
           ..Awarded) %>% 
  select(-c(Applicant.ACE.Area,
            Constituency)) %>% 
  relocate(Local.Authority) %>% 
  relocate(Region.ONS, .after = Grant) %>% 
  datatable(options = list(pageLength = 5),
            colnames = c("Local Authority", 
                         "Organisation", 
                         "Total £ awarded",
                         "Region",
                         "Discipline",
                         "Scheme",
                         "NPO?"),
            rownames = FALSE,
            filter = "top") %>%
  formatCurrency('Grant',
                 currency = "£", 
                 interval = 3, 
                 mark = ",",
                 digits = 0) %>% 
  formatStyle("Grant")
```

## How much money did NPOs and non-NPOs get by local authority?

We can (briefly!) extend this by getting the total amount of money going to each local authority that went to NPOs, and that went to non-NPOs. This table also tells you what % of the money in each LA went to NPOs.

```{r}
crf_money %>% 
  group_by(Local.Authority,
           is_npo) %>% 
  summarise(total_money = 
              sum(..Awarded)) %>% 
  pivot_wider(names_from = is_npo,
              values_from = total_money,
              names_repair = "universal") %>% 
  mutate(NPO = replace_na(NPO, 0)) %>% 
  mutate(Not.NPO = replace_na(Not.NPO, 0)) %>% 
  mutate(percent_npo = 
           round(NPO*100/(Not.NPO+NPO)) )%>% 
  datatable(options = list(pageLength = 5),
            colnames = c("Local Authority", 
                         "Money to non-NPOs", 
                         "Money to NPOs",
                         "% of money to NPOs"),
            rownames = FALSE,
            filter = "top") %>%
  formatCurrency('Not.NPO',
                 currency = "£", 
                 interval = 3, 
                 mark = ",",
                 digits = 0) %>% 
  formatStyle("Not.NPO") %>%
  formatCurrency('NPO',
                 currency = "£", 
                 interval = 3, 
                 mark = ",",
                 digits = 0) %>% 
  formatStyle("NPO")

```